---
title: "P8105 Homework 6"
author: "Ruiyang Li"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

This is my solution to HW6.

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)             # for stepAIC()
library(modelr)
library(purrr)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


### Problem 1

```{r homicide_df,collapse=TRUE}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  dplyr::select(city_state, resolution, victim_age, victim_race, victim_sex)
```


Start with one city: Baltimore, MD.

```{r baltimore_rlts,collapse=TRUE}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  dplyr::select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```


Try this across cities.

```{r all_cities_rslts,collapse=TRUE}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  dplyr::select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  dplyr::select(city_state, term, OR, starts_with("CI")) 
```


Create a plot that shows the estimated ORs and CIs for each city. 

```{r or_plt,collapse=TRUE}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

Load and clean the data for regression analysis. 

* Convert numeric *factor* variables back to factor. 
* There is no missing data observed. 

```{r baby_df,collapse=TRUE}
# load and clean data
baby_df = 
  read_csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = factor(babysex), 
    frace = factor(frace), 
    malform = factor(malform), 
    mrace = factor(mrace)
    )

# check missing values
baby_df %>%  summarise(na_count = sum(is.na(.))) # total # of na's
```

Propose a regression model for birthweight. 

* We can use step-wise selection procedure to select our regression model. 
* The step-wise selection procedure iteratively adds and/or removes important predictor based on AIC score. 
* Based on the results from step-wise selection, my proposed model is: `bwt` = `β_0` + `β_1` `babysex2` + `β_2` `bhead` + `β_3` `blength` + `β_4` `delwt` + `β_5` `fincome` + `β_6` `gaweeks` + `β_7` `mheight` + `β_8` `mrace2` + `β_9` `mrace3` + `β_10` `mrace4`  + `β_11` `parity` + `β_12` `ppwt` + `β_13` `smoken`

```{r stepwise_reg,collapse=TRUE}
# fit the full model 
full_mod = lm(bwt ~ ., data = baby_df)

# stepwise regression model
step_mod = stepAIC(full_mod, direction = "both", trace = F)
step_mod %>% broom::tidy()

# selected model 
step_mod$call
step_mod$anova
```

* Below is the plot of model residuals vs. fitted values. 

```{r plt_resid_vs_pred_step_mod}
# fit selected model 
fit = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken, 
         data = baby_df)
  
# plot resid vs. pred
baby_df %>% 
  add_predictions(fit) %>% 
  add_residuals(fit) %>% 
  ggplot(aes(y = resid, x = pred)) + 
  geom_point(alpha = .2) + 
  labs(
    title = "Plot of model residuals vs. fitted values",
    subtitle = "Stepwise model",
    x = "Fitted Values",
    y = "Residuals"
  )
```


Compare the selected model to
 
* Main effect model: using length at birth and gestational age as predictors
* Interaction model: using head circumference, length, sex, and all interactions as predictors

Results: 

* From the violin plot below, we can see that our selected model performed the best among the three models as it has the lowest rmse. 
* The interaction model performed the second best. 
* The main effect model performed the worst among the three. 

```{r compare_models,collapse=TRUE}
# get cv training and testing datasets
set.seed(20201209)
cv_df = 
  crossv_mc(baby_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

# fit models and get RMSEs
cv_df = 
  cv_df %>% 
  mutate(
    main_mod = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit_mod = map(.x = train, 
                  ~lm(bwt ~ babysex + bhead + blength + delwt + fincome +
                        gaweeks + mheight + mrace + parity + ppwt + smoken, 
                      data = .x)),
    int_mod = map(.x = train, ~lm(bwt ~ bhead:blength:babysex, data = .x))
  ) %>% 
  mutate(
    rmse_main = map2_dbl(.x = main_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_fit = map2_dbl(.x = fit_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_int = map2_dbl(.x = int_mod, .y = test, ~rmse(model = .x, data = .y))
  )

# compare models in terms of the cv prediction error (rmse)
cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() + 
  labs(
    title = "Violin plot",
    subtitle = "fit: step-wise model; int: interaction model; main: main effect model")
```



## Problem 3

Download data. 

```{r weather_df,collapse=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  dplyr::select(name, id, everything())
```

Regression of interest: 

* tmax = β_0 + β_1 tmin

Goal: 

* To use 5000 bootstrap samples to learn about the distribution of r^2 and log(hat β_0 ∗ hat β_1) 


For each bootstrap sample, produce estimates of these two quantities. 

```{r btsrp_rslt_est,collapse=TRUE}
# create 5000 bootstrap samples
set.seed(20201209)
bootstrap_rslt = 
  weather_df %>% 
  bootstrap(5000, id = "strap_number") %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)), 
    r_squared = map(models, broom::glance),
    results = map(models, broom::tidy)
  )

# produce estimates of r^2 + log(\hat(\beta_0) * \hat(\beta_1))
bootstrap_est = 
  bootstrap_rslt %>% 
  dplyr::select(strap_number, r_squared, results) %>% 
  unnest(r_squared) %>% 
  dplyr::select(strap_number, r.squared, results) %>% 
  unnest(results) %>% 
  janitor::clean_names() %>%  
  dplyr::select(strap_number, r_squared, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate
  ) %>% 
  rename(beta_0 = `(Intercept)`, beta_1 = tmin) %>% 
  mutate(log_beta0beta1 = log(beta_0 * beta_1)) %>% 
  dplyr::select(-starts_with("beta_"))

head(bootstrap_est)
```

Plot the distribution of these estimates. 

* r^2 has a bell-shaped distribution, with mean at `r round(mean(bootstrap_est$r_squared), 3)`. It ranges from `r round(min(bootstrap_est$r_squared), 3)` to `r round(max(bootstrap_est$r_squared), 3)` 
* log(hat β_0 ∗ hat β_1) also has a bell-shaped distribution, with mean at `r round(mean(bootstrap_est$log_beta0beta1), 3)`. It ranges from `r round(min(bootstrap_est$log_beta0beta1), 3)` to `r round(max(bootstrap_est$log_beta0beta1), 3)`.  

```{r distn_plt,collapse=TRUE}
# plot distribution of r^2
r2_distn = 
  bootstrap_est %>% 
  ggplot(aes(x = r_squared)) + 
  geom_density() + 
  labs(
    title = "Distribution of r^2",
    subtitle = "Based on 5000 bootstrap samples", 
    x = "R squared")

# plot distribution of log(\hat(\beta_0) * \hat(\beta_1))
log_distn = 
  bootstrap_est %>% 
  ggplot(aes(x = log_beta0beta1)) + 
  geom_density() + 
  labs(
    title = "Distribution of log(hat β_0 ∗ hat β_1)",
    subtitle = "Based on 5000 bootstrap samples", 
    x = "log(hat β_0 ∗ hat β_1)")

r2_distn + log_distn
```

Using these 5000 estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(hat β_0 ∗ hat β_1). 

* The 95% confidence interval for r^2 is (0.894, 0.927).
* The 95% confidence interval for log(hat β_0 ∗ hat β_1) is (1.965, 2.059).
 
```{r 95ci,collapse=TRUE}
# construct bootstrap CI
bootstrap_est %>% 
  dplyr::summarize(
    r2_ci_lower = quantile(r_squared, 0.025),
    r2_ci_upper = quantile(r_squared, 0.975), 
    log_ci_lower = quantile(log_beta0beta1, 0.025),
    log_ci_upper = quantile(log_beta0beta1, 0.975)
  ) %>% 
  knitr::kable(digits = 3) 
```


